---
title: "Regression"
output: html_document
date: "2024-07-20"
---

```{r}
# imports
require(ggplot2)
require(zoo)
require(reshape2)
require(fitdistrplus)
require(lmtest)
```


## Load and pre-process data
Assumption: the three csv files were downloaded into data/, i.e. these files exist
data/DGS5.csv 
data/DGS10.csv 
data/T10YIE.csv


```{r}
# read data, set types, encode missings
data_path <- "data/"
inflat <- read.csv(
  file=paste0(data_path, "T10YIE.csv"), 
  colClasses = c("DATE"="Date", "T10YIE"="numeric"),
  na.strings = c(".", "", "NA")
)
yield5 <- read.csv(
  file=paste0(data_path, "DGS5.csv"),
  colClasses = c("DATE"="Date", "DGS5"="numeric"),
  na.strings = c(".", "", "NA")
)
yield10 <- read.csv(
  file=paste0(data_path, "DGS10.csv"),
  colClasses = c("DATE"="Date", "DGS10"="numeric"),
  na.strings = c(".", "", "NA")
)
```


## explore basic location and dispersion statistics
```{r}
summary(inflat)
```
```{r}
summary(yield5)
```
```{r}
summary(yield10)
```


## fill missings by carrying last observation forward
```{r}
inflat <- na.locf(inflat)
yield5 <- na.locf(yield5)
yield10 <- na.locf(yield10)

stopifnot(sum(is.na(inflat))==0)
stopifnot(sum(is.na(yield5))==0)
stopifnot(sum(is.na(yield10))==0)
```

## check whether all have same dates
inflat has one day more, 2024-07-19 (we can omit this date to align date axis)

```{r}
df.list <- list(inflat=inflat, yield5=yield5, yield10=yield10)
df.pairs <- combn(c("inflat", "yield5", "yield10"), 2)
for (i in 1:dim(df.pairs)[2]){
  name1 <- df.pairs[1,i]
  name2 <- df.pairs[2,i]
  
  df1 <- df.list[[name1]]
  df2 <- df.list[[name2]]
  
  date_setdiff <- setdiff(unique(df1$DATE), unique(df2$DATE))
  if(length(date_setdiff)>0){
    print(paste("dates set diff between", name1, "-", name2, ":", as.Date(date_setdiff)))
  }
  
  date_setdiff_back <- setdiff(unique(df2$DATE), unique(df1$DATE))
  if(length(date_setdiff_back)>0){
    print(paste("dates set diff between", name2, "-", name1, ":", as.Date(date_setdiff_back)))
  }
}  
```
## put all in one dataframe to align dates

```{r}
df <- merge(merge(inflat, yield5), yield10)
stopifnot(length(unique(df$DATE))==length(unique(yield5$DATE)))
stopifnot(sum(is.na(df))==0)
```


```{r}
# save dataframe
save(df, file="dataframes/reg.Rda")
```


```{r}
# plot all time series at once
df.melt <- melt(df, id="DATE", measure.vars=c("T10YIE", "DGS5", "DGS10"), variable.name="TICKER", value.name="RATE")
ggplot(df.melt, aes(x=DATE, y=RATE, color=TICKER)) + geom_line()
```


## evaluate assumptions of linear regression (a priori sanity checks)
assumption: the response is DGS10 which we want to regress on the explanatory variable T10YIE

linearity: from the plot it is doubtful whether the relationship between x=T10YIE and E[DGS10] is linear

```{r}
ggplot(df, aes(x=T10YIE, y=DGS10)) + geom_point()
```

independence violated: observations of dependent and independent variable are highly autocorrelated
```{r}
acf(df$T10YIE)
acf(df$DGS10)
```

normality of response: Shapiro test rejects the H0 of normality
```{r}
shapiro.test(df$DGS10) # p-value < 2.2e-16
```

normal distribution of response
- deviations from normal distribution, especially in tails
- don't need this assumption for prediction
```{r}
fit.norm <- fitdist(df$DGS10, distr="norm")
plot(fit.norm)
```


## regress 10y yield (y) against inflation (x)
```{r}
fit <- lm(DGS10 ~ 1 + T10YIE, data=df)
summary(fit)
```


```{r}
ggplot(df, aes(x=T10YIE, y=DGS10)) + 
  geom_point() +
  geom_line(aes(x=T10YIE, y=fit$fitted.values), color="red")
```

# diagnose residual


hist shows heavy tailed residuals
```{r}
hist(fit$residuals)
```


Tuckey-Anscombe plot: 
- assumptions of zero mean and homoscedasticity are violated!
- shape of mean curve could suggest a missing cubic term
(https://www.londschien.ch/teaching/stat-modelling/2022-11-14)
```{r}
plot(fit, which=1)
```


Q-Q plot shows deviation from normal in tails
```{r}

plot(fit, which=2)

```
