---
title: "Yield Curve Analysis"
output: html_document
date: "2024-07-19"
---

```{r}
# imports
library(rvest)
library(xml2)
library(readxl)
library(ggplot2)
library(viridis)
```

## Set paths

```{r}
rfr_data_path <- "data/rfr/"

# check data dir existence and set data download flag
if(dir.exists(rfr_data_path)){
  download_data = F
}else{
  download_data = T
  # create data directory
  dir.create(rfr_data_path, recursive = T)
}

# checkpoint directory
df_path <- "dataframes/"
df_rfr_path <- paste0(df_path, "rfr.Rda")

```

## Download data or load checkpoint

```{r}
if(download_data){
  ## get html contents
  url <- "https://www.eiopa.europa.eu/tools-and-data/risk-free-interest-rate-term-structures_en#risk-free-rates-previous-releases-and-preparatory-phase"
  webpage <- read_html(url)
  
  div_ids <- c(
    "paragraph-165-0-content", # 2024
    "paragraph-165-1-content"  # 2023
  )
  
  ## get all download links
  download_links <- vector(mode="character")
  for (div_id in div_ids){
    section <- html_node(webpage, css = paste0("#", div_id))  # Replace with the correct CSS selector
    links <- html_nodes(section, "a")
    download_links <- c(download_links, html_attr(links, "href"))
  }
  
  ## download all files
  base_url <- "https://www.eiopa.europa.eu/"
  for (file_link in download_links){
    file_name <- strsplit(file_link, "filename=")[[1]][2]
    download.file(
      paste0(base_url, file_link),
      paste0(rfr_data_path, file_name),
      mode = "wb"
    )
  }
  
  # check whether names are correct - if not rename file manually
  for (file in list.files(rfr_data_path)){
    if(!startsWith(file, "EIOPA_RFR_") | !endsWith(file, ".zip")){
      print(paste(file, "does not comply with naming convention."))
    }
  }
  
  # check whether number of files is plausible
  n_dates <- 1 + 12 + 6
  stopifnot(length(list.files(rfr_data_path)) == n_dates) # 2022 + 2023 + 2024
  
  
  ## Prepare dataframe
  
  # unzip all files
  for (file in list.files(rfr_data_path, full.names = T)){
    unzip(file, exdir=rfr_data_path)
  }
  
  # extract country and tenor wise interest rates (RFR without volatility adjustment)
  
  process_rfr_df <- function(df, file_name){
    # remove rows not needed
    df <- df[-c(1:8),]
    row.names(df) <- NULL
    colnames(df)[1] <- "Tenor"
  
    # date column
    date_str <- strsplit(file_name, "_")[[1]][3]
    df[,"Date"] <- as.Date(date_str, format="%Y%m%d")
  
    # put date column first
    df <- df[,c(ncol(df), 1:ncol(df)-1)]
  
    return(df)
  }
  
  # load excels, extract sheet, process sheet and concat to one comprehensive data frame
  df <- data.frame()
  for (file_name in list.files(rfr_data_path, full.names = T)){
    if(grepl("_Term_Structures.xlsx", file_name)){
      print(paste("Loading", basename(file_name)))
      df_sub <- data.frame(
        read_excel(file_name, sheet="RFR_spot_no_VA")
      )
      base_file_name <- basename(file_name)
      print(paste("Processing", base_file_name))
      df_sub <- process_rfr_df(df_sub, base_file_name)
  
      df <- rbind(df, df_sub)
    }
  }
  
  # set data types
  df$Tenor <- as.integer(df$Tenor)
  cols.rates <- c(3:dim(df)[2])
  df[cols.rates] <- sapply(df[cols.rates], as.double)
  
  # check if all files were concatenated
  stopifnot(length(unique(df$Date)) == n_dates)
  # check if all Tenors are there
  stopifnot(nrow(df) == n_dates * length(unique(df$Tenor)))
  # check if we have NAs for the locations
  stopifnot(sum(is.na(df))==0)
  # check if RFRs make sense
  stopifnot(all(df[,cols.rates]>-0.01) & all(df[,cols.rates] < 1))
  
  # check extreme values
  # df[which(df<0, arr.ind = T)] # check negative yields
  # max(df[cols.rates])
  
  # save dataframe
  save(df, file=df_rfr_path)
  
} else {
  # Load checkpoint
  load(df_rfr_path)
}
```


## Visualize the changes in US and Euro yield curve across tenors since 2022
Assumption: changes in yield curve = yield curve

```{r}
plot_yield_curves <- function(region){
  ggplot(df, aes(x=Tenor, y=df[,region], color=as.factor(Date))) + 
    geom_line(aes(color = as.factor(Date))) + 
    scale_color_viridis_d() +
    labs(title = paste(region, "Yield Curves"),
         x = "Maturity (Years)",
         y = "Yield",
         color = "Date") +
    theme_minimal() +
    theme(legend.position = "bottom")
}
```

```{r}
plot_yield_curves("Euro")
```

```{r}
plot_yield_curves("United.States")
```

## Compute and visualise the correlation between US/Euro/UK/China rates focusing only on 2/5/10/20/30y tenor points

Assumption: correlation between ALL given region pairs should be calculated for specific tenors across ALL dates 

```{r}
# computation of correlation data frame
tenors <- c(2,5,10,20,30)
regions <- c("United.States", "Euro", "United.Kingdom", "China")


region.pairs <- combn(regions,2)
df_cors <- data.frame()

for (i in 1:dim(region.pairs)[2]){
  region1 <- region.pairs[1,i]
  region2 <- region.pairs[2,i]
  
  name.region.pair <- paste(region1, "/", region2)
  
  for(j in 1:length(tenors)){
    tenor <- tenors[j]
    rfr1 = df[df$Tenor==tenor, region1]
    rfr2 = df[df$Tenor==tenor, region2]
    cor_region_tenor <- cor(rfr1, rfr2)
    df_cors <- rbind(df_cors, c(name.region.pair, tenor, cor_region_tenor))
  }
}
colnames(df_cors) <- c("Region.Pair", "Tenor", "Correlation")
df_cors$Correlation <- as.numeric(df_cors$Correlation)
df_cors$Tenor <- as.numeric(df_cors$Tenor)

head(df_cors)
```


```{r}
## visualization of correlations across tenors
ggplot(df_cors, aes(x=Tenor, y=Correlation, color=as.factor(Region.Pair))) + 
  geom_line(aes(color = as.factor(Region.Pair)), size = 1) + 
  scale_color_viridis_d() +
  scale_x_continuous("Maturity (Years)", labels = as.character(tenors), breaks = tenors) +
  labs(title = "RFR Correlation among Region Pairs",
       y = "Correlation",
       color = "Region Pair") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Compute and visualise 2 year-10 year and 2 year-30 year slope for US and Euro rates

Assumption: ay-by slope means slope of straight line between (b, yield(b)) and (a, yield(a))
Assumption: change of slope should be visualized over time (insted of slope distribution)

```{r}
# slope computation
df_slopes <- data.frame()

get_region_tenor_df_slope <- function(region, tenor.start, tenor.end){
  rfr_start <- df[df$Tenor==tenor.start, region]
  rfr_end <- df[df$Tenor==tenor.end, region]
  tenor.diff <- tenor.end - tenor.start
  rfr_slopes <- (rfr_end - rfr_start) / tenor.diff
  dates <- unique(df$Date)
  region_tenor_df_slope <- data.frame(
    Region=rep(region, length(dates)),
    Tenor.Pair=rep(paste0(tenor.start, "y", "-", tenor.end, "y"), length(dates)),
    Date=dates,
    RFR.Slope=rfr_slopes
  )
  return(region_tenor_df_slope) 
}
regions <- c("Euro", "United.States")
for(region in c("Euro", "United.States")){
  for (tenor.pair in list(c(2,10), c(2,30))){
    region_tenor_df_slope <- get_region_tenor_df_slope(
      region, tenor.pair[1], tenor.pair[2]
    )
    df_slopes <- rbind(df_slopes, region_tenor_df_slope)
  }
}
stopifnot(dim(df_slopes)[1] == length(unique(df$Date))*2*2)
head(df_slopes)
```

```{r}
# slope plotting
plot_rfr_slope <- function(region){
  ggplot(df_slopes[df_slopes$Region==region,], aes(x=Date, y=RFR.Slope, color=as.factor(Tenor.Pair))) + 
    geom_line(aes(color = as.factor(Tenor.Pair)), size = 1) + 
    scale_color_viridis_d() +
    labs(title = paste(region, "RFR Slopes"),
         color = "Tenor.Pair") +
    theme_minimal() +
    theme(legend.position = "bottom")
}
```


```{r}
plot_rfr_slope("Euro")
```

```{r}
plot_rfr_slope("United.States")
```



